---
title: "Getting started with the networkinference R package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with the networkinference R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

## Necessary packages for this vignette

```{r setup}
library(networkinference) # Package implementing the methods of this paper
library(nett) # For spectral clustering
library(RSpectra) # Dependency of nett package

set.seed(1234) # Reproducibility
```

# Examples

## Networks with Gaussian edges

Suppose we observe a network $n = 60$ nodes. In the Gaussian setting, the edges encoded in the adjacency matrix obey $A_{ij} \overset{\text{ind.}}{\sim} N(M_{ij}, \tau^2)$, where we treat $M_{ij}$ as unknown and $\tau^2$ as known. Let us simulate an adjacency matrix from a setting where there are three "true communities," where the intra-community connectivity is $\rho_1 = 25$, the inter-community connectivity is $\rho_2 = 18$, and the known edge variance is $\tau^2 = 4^2$.

```{r}
# Some global information
n <- 60
n_per_community <- 20
tau <- 4 # Known standard deviation of the edges
rho_1 <- 25 # Intra-community connectivity
rho_2 <- 18 # Inter-community connectivity

# Build tiles which we will use to construct our mean matrix M
intra_comm <- matrix(rep(rho_1, n_per_community^2), nrow = n_per_community)
inter_comm <- matrix(rep(rho_2, n_per_community^2), nrow = n_per_community)

# Construct mean matrix
M <- rbind(cbind(intra_comm, inter_comm, inter_comm),
           cbind(inter_comm, intra_comm, inter_comm),
           cbind(inter_comm, inter_comm, intra_comm))

# Draw a network using this mean matrix
A <- matrix(rnorm(n = n^2, mean = as.vector(M), sd = tau), nrow = n)
```

Visualizing the adjacency matrix with an image (rows run from top to bottom, columns run from left to right), we can see that there are a greater density of connections within communities than there are between communities.

```{r, fig.width = 4, fig.height = 4}
par(mar = c(0.5, 0.5, 0.5, 0.5))
par(pty = "s")
image(t(A[nrow(A):1,]), axes = FALSE, asp = 1)
```

Now, let's use the `split_network()` function from the package to split this adjacency matrix $A$ into our train and test adjacency matrices. As arguments, we must provide our distributional setting (Gaussian), our splitting parameter $\epsilon = 0.5$, as well as our known variance $\tau^2 = 4^2$.

```{r, fig.width=6, fig.height = 3}
gaussian_split_results <- split_network(A, distribution = "gaussian", 
                                        epsilon = 0.5, tau = tau)
Atr <- gaussian_split_results$Atr
Ate <- gaussian_split_results$Ate

# Visualize both the Atr and Ate matrices
layout(matrix(1, 2, nrow = 1))
par(mfrow = c(1, 2), mar = c(1, 1, 1, 1))
image(t(Atr[nrow(Atr):1,]), axes = FALSE, asp = 1)
title("Train network")
image(t(Ate[nrow(Ate):1,]), axes = FALSE, asp = 1)
title("Test network")
```

Now, let us conduct community estimation using `Atr` using $K=4$ communities. (Note that this is a misspecified number of communities!)

```{r}
estimated_communities <- nett::spec_clust(Atr, K = 4)
estimated_communities # A vector specifying each node's estimated community
```

Let's now try to conduct inference for the **mean intra-community connectivity within the 1st estimated community.** To do this, we will want to set an appropriate linear combination vector $u$ which is of length $K^2$, where $K$ is the number of estimated communities.

```{r}
u_vector <- c(1, 0, 0, 0, 
              0, 0, 0, 0, 
              0, 0, 0, 0,
              0, 0, 0, 0)

# Conduct inference for the selected target (mean connectivity within the
# first estimated community)
gaussian_inference <-
  infer_network(Ate = Ate, u = u_vector,
                communities = estimated_communities,
                distribution = "gaussian",
                epsilon = 0.5, K = 4, tau = tau)
```

The resulting `gaussian_inference` contains two pieces, the estimate `gaussian_inference$estimate` ($\hat{\theta}(A^{(\text{tr})})$ in the accompanying paper), as well as the estimate of the variance `gaussian_inference$estimate_variance` ($\hat{\sigma}^2$ in the accompanying paper). In this case, we find that

$$\hat{\theta}(A^{(\text{tr})}) = `r round(gaussian_inference$estimate, 4)`, \quad \hat{\sigma}^2 = `r round(gaussian_inference$estimate_variance, 4)`.$$

So, we can construct a 90\% confidence interval for this target with the following:

```{r}
alpha <- 0.10
margin_of_error <- qnorm(1 - alpha/2) * sqrt(gaussian_inference$estimate_variance)
ci_lower_bound <- gaussian_inference$estimate - margin_of_error
ci_upper_bound <- gaussian_inference$estimate + margin_of_error

print(paste0("Estimate of selected target is ", round(gaussian_inference$estimate, 4)))
print(paste0("Confidence interval for selected target is [",
             round(ci_lower_bound, 4), ", ", round(ci_upper_bound, 4), "]"))
```

Because we simulated the data ourselves, we can check whether this confidence interval actually contained the true target of inference.

```{r}
true_target <- check_target_of_inference(M = M, u = u_vector,
                                         communities = estimated_communities,
                                         K = 4)
true_target_contained <- (ci_lower_bound <= true_target) & 
  (true_target <= ci_upper_bound)

# Print results
print(paste0("True target of inference is ", round(true_target, 4)))
print("Did the 90% confidence interval contain the true target of inference?")
print(paste0("-> ", as.character(true_target_contained)))

```

## Networks with Poisson edges

The workflow for networks with Poisson edges is nearly identical to that of Gaussian edges, so a concrete example is omitted. The only difference for networks with Poisson edges is that in the `split_network()` and `infer_network()` functions, you should set `distribution = "poisson"`. Note also that the `tau` parameter is unnecessary and will be ignored.

## Networks with Bernoulli edges

Once again, we consider a setting with $n=60$ nodes. In the Bernoulli setting, we assume $A_{ij} \overset{\text{ind.}}{\sim} \text{Bernoulli}(M_{ij})$, where $M_{ij}$ is unknown. Let us simulate in a setting where there are three "true communities," where the intra-community connectivity is $\rho_1 = 0.80$, and the inter-community connectivity is $\rho_2 = 0.40$.

```{r}
n <- 60
n_per_community <- 20
rho_1 <- 0.80
rho_2 <- 0.40

# Used to construct the mean matrix
intra_comm <- matrix(rep(rho_1, n_per_community^2), nrow = n_per_community)
inter_comm <- matrix(rep(rho_2, n_per_community^2), nrow = n_per_community)

# Construct mean matrix
M <- rbind(cbind(intra_comm, inter_comm, inter_comm),
           cbind(inter_comm, intra_comm, inter_comm),
           cbind(inter_comm, inter_comm, intra_comm))

# Draw a network using this mean matrix
A <- matrix(rbinom(n = n^2, size = 1, prob = as.vector(M)), nrow = n)
```

Visualizing the adjacency matrix with an image (rows run from top to bottom, columns run from left to right), we can see that there are a greater density of connections within communities than there are between communities. In contrast to the Gaussian adjacency matrix from before, the edge connections are unweighted (binary) rather than weighted (continuous or integer-valued).

```{r}
par(mar = c(0.5, 0.5, 0.5, 0.5))
par(pty = "s")
image(t(A[nrow(A):1,]), axes = FALSE, asp = 1)
```

Now, we can use the `split_network()` function from the package to split this into our train and test adjacency matrices. To split with a parameter $\gamma = 0.15$, we will pass in `gamma = 0.15` into the function, and specify our distributional setting.

```{r, fig.width=6, fig.height = 3}
bernoulli_split_results <- split_network(A, distribution = "bernoulli",
                                         gamma = 0.15)
Atr <- bernoulli_split_results$Atr
Ate <- bernoulli_split_results$Ate

# Visualize both the Atr and Ate matrices
layout(matrix(1, 2, nrow = 1))
par(mfrow = c(1, 2), mar = c(1, 1, 1, 1))
image(t(Atr[nrow(Atr):1,]), axes = FALSE, asp = 1)
title("Train network")
image(t(Ate[nrow(Ate):1,]), axes = FALSE, asp = 1)
title("Test network")
```

Now, let us conduct community estimation using `Atr` using $K=4$ communities. (Note that this is a misspecified number of communities!)

```{r}
estimated_communities <- nett::spec_clust(Atr, K = 4)
estimated_communities # A vector specifying each node's estimated community
```

Let’s now try to conduct inference for the **mean intra-community connectivity within the 1st estimated community.** To do this, we will want to set an appropriate linear combination vector $u$ which is of length $K^2$, where $K$ is the number of estimated communities.

```{r}
u_vector <- c(1, 0, 0, 0, 
              0, 0, 0, 0, 
              0, 0, 0, 0,
              0, 0, 0, 0)

# Conduct inference for the selected target (mean connectivity within the
# first estimated community)
bernoulli_inference <-
  infer_network(Ate = Ate, u = u_vector,
                communities = estimated_communities,
                distribution = "bernoulli",
                Atr = Atr,
                gamma = 0.15, K = 4)
```

The resulting object `bernoulli_inference` contains two pieces, the estimate `bernoulli_inference$estimate` ($\hat{\xi}(A^{(\text{tr})})$ in the paper), as well as the estimate of the variance `bernoulli_inference$estimate_variance` ($\hat{\sigma}^2$ in the paper). In this case, we find that

$$\hat{\xi}(A^{(\text{tr})}) = `r round(bernoulli_inference$estimate, 4)`, \quad \hat{\sigma}^2 = `r round(bernoulli_inference$estimate_variance, 4)`.$$

Thus, we construct a 90\% confidence interval with the following:

```{r}
alpha <- 0.10
margin_of_error <- qnorm(1 - alpha/2) * sqrt(bernoulli_inference$estimate_variance)
ci_lower_bound <- bernoulli_inference$estimate - margin_of_error
ci_upper_bound <- bernoulli_inference$estimate + margin_of_error

print(paste0("Confidence interval for selected target is [",
             round(ci_lower_bound, 4), ", ", round(ci_upper_bound, 4), "]"))
```

Because we simulated the data ourselves, we can check how this matches up with the true targets of interest, both $\theta(A^{(\text{tr})})$ and $\xi(A^{(\text{tr})})$. Please see the accompanying paper for a reminder on the differences between these two targets.

```{r}
true_target_theta <- check_target_of_inference(M = M, u = u_vector,
                                               communities = estimated_communities,
                                               K = 4)
true_target_xi <- check_target_of_inference(M = M, u = u_vector,
                                            communities = estimated_communities,
                                            K = 4, bernoulli_target = TRUE,
                                            gamma = 0.15, Atr = Atr)
```

In this case, we find that

$$\theta(A^{(\text{tr})}) = `r round(true_target_theta, 4)`, \quad \xi(A^{(\text{tr})}) = `r round(true_target_xi, 4)`.$$
